---
title: "Portfolio Optimization"
author: "Navein Suresh, Suryaveer Singh"
date: "~"
format: html
editor: visual
---

## Overview

Our objective in this project is to optimize a portfolio of stocks for risk-adjusted returns. We will be focusing on a 5-stock portfolio consisting of the leading semiconductor companies in the market:

1.  *Advanced Micro Device, Inc. (NASDAQ: AMD)*
2.  *Intel Corp. (NASDAQ: INTC)*
3.  *NVIDIA Corp. (NASDAQ: NVDA)*
4.  *Super Micro Computer Inc. (NASDAQ: SMCI)*
5.  *Taiwan Semiconductor Mfg. Co. Ltd. (NYSE: TSM)*

We make use of various financial models, which we go into depth later on while also leverage the tidyquant package on R as our convenient tool for portfolio optimization. Using historical data, we will test our portfolio’s performance and compare it to the performance of two benchmark index funds — S&P 500 and NASDAQ.

Be sure to have to following libraries installed and imported before proceeding:

```{r}
library(tidyquant)
library(dplyr)
library(ggplot2)

```

Our testing time period will be from January 1, 2000 to January 1, 2024. This allows us to account for a range of economic trends — downswings such as the 2008 financial crisis and the zero interest rate growth period during the global COVID-19 pandemic.

We begin with an equally weighted portfolio of the above stocks, each at 20% equity holding. For capital consideration, we will assume we start with \$10,000 (USD).

```{r}
stock_tickers <- c("AMD", "INTC", "NVDA", "SMCI", "TSM")
weights <- rep(0.2, 5)
initial_investment <- 10000
ten_year_treasury_yield_table = arrange(tq_get("DGS10", get = "economic.data"), desc(date))
latest_ten_year_treasury_yield = slice(ten_year_treasury_yield_table, 1)$price

stock_allocations <- weights * initial_investment

portfolio <- data.frame("Stock" = stock_tickers, "Investment ($)" = stock_allocations)
print(latest_ten_year_treasury_yield)
print(portfolio)

```

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).
